Semantic Textual Similarity Using SVM and Deep Learning

### 2017-07-06 paper/main.tex
Fixed almost all mandatory fixes

\title{Semantic Textual Similarity using deep learning}

@@

\address{
University of Zagreb, Faculty of Electrical Engineering and Computing\\
Unska 3, 10000 Zagreb, Croatia\\
\texttt{\{bruno.gavranovic,neven.miculinic,stipan.mikulic\}@fer.hr}\\
}

@@

% \begin{table}[h!]
% \caption{STS score description, adapted from~\citep{agirre2016semeval}}
% \label{tab:sts-score}
% \begin{center}
% 	\resizebox{\columnwidth}{!}{
% \begin{tabularx}{\columnwidth}{llr}
% \toprule
% Score & Explanation\\
% \midrule
% 5 & \textit{Two sentences are completely equivalent}\\
% & The bird is bathing in the sink\\
% & Birdie is washing itself in the water basin.\\
% \midrule
% 4 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
%   & \multicolumn{1}{m{6.5cm}}{In May 2010, the troops attempted to invade
% Kabul.} \\
% & \multicolumn{1}{m{6.5cm}}{The US army invaded Kabul on May 7th last
% year, 2010.} \\
% \midrule
% 3 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
%   & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
% suspect.}\\
% & ``He is not a suspect anymore.'' John said.\\
% \midrule
% 2 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
% &They flew out of the nest in groups. \\
% &They flew into the nest together. \\
% \midrule
% 1 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
% & The woman is playing the violin.\\
% & The young lady enjoys listening to the guitar.\\
% \midrule
% 0 & \textit{The two sentences are completely dissimilar.}\\
%   & \multicolumn{1}{m{6.5cm}}{John went horse back riding at dawn with a whole group of friends.}\\
%   & \multicolumn{1}{m{6.5cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
% \bottomrule
% \end{tabularx}
% }
% \end{center}
% \end{table*}
% 
\section{Related work}

@@

\section{Extent of the paper}

@@

\subsection{Data analysis}
We used dataset from SemEval 2017
Task 1\footnote{http://alt.qcri.org/semeval2017/task1/}.
The dataset has 250 instances. Each instance consists of two sentences and our tash is to predict semantic similarity between them. In order to get some deeper insights on data we did some minor data analysis. We got to conclusion that we are dealing with very short text. Also, we concluded that the smaller difference between lengths of sentences is, the more similar they are. This can be seen clearer on following plots.

\begin{table}
\caption{Stats about lengths of sentences}
\label{tab:narrow-table}
\begin{center}
\begin{tabular}{cll}
\toprule
& Without stopwords & With stopwords \\
\midrule
Sentence length & 33.004 & 43.838 \\
Tokens in sentence & 5.488 & 8.702 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{sen_len_no_stop.png}
%\caption{Sentence lengths without stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{sen_len_with_stop.png}
%\caption{Sentence lengths with stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{tokens_no_stop.png}
%\caption{Number of tokens without stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{tokens_with_stop.png}
%\caption{Number of tokens with stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
\subsection{Baseline models}

@@

\subsubsection{Jaccard similarity baseline}

@@

\subsubsection{Cosine similarity}

@@

\begin{table}
\caption{Averaged scores}
\label{tab:narrow-table}
\begin{center}
\begin{tabular}{ccc}
\toprule
& R2 & Pearson\\
\midrule
score & 0.5898 & 0.7795 \\
std & 0.011393 & 0.006625 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\subsection{Deep learning}

@@

\subsubsection{Recurrent networks}

@@

\begin{table}
\caption{Sample network output}
\label{tab:sample_output}
\begin{center}
\begin{tabular}{c}
\toprule
Sample output of the LSTM \\
\midrule
\multicolumn{1}{m{6.5cm}}{Little Joe Otter had had no honour pointing to help him so that they
heard of his own fish.  And it was for the beginning of the Little Fir Tree and
scrambled to the sheep. 
Come, let us go and let the right out of this tree, and the LORD had chanced
to the next tabernacce of Mischians, when the king was carried the short of the children of the Green.}
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

\subsection{All results}

@@

\begin{table}
	\caption{Model comparison}
\label{tab:all_results}
\begin{center}
\begin{tabular}{ccc}
\toprule
Model & Pearson coefficient \\
\midrule
Jaccard similarity & 0.73 \\
Cosine baseline & 0.73 \\
LSTM & 0.61 \\
SVM & 0.78 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

>>>>

\title{Semantic Textual Similarity Using SVM and Deep Learning}

@@

\address{
University of Zagreb, Faculty of Electrical Engineering and Computing\\
Unska 3, 10000 Zagreb, Croatia\\
\texttt{\{bruno.gavranovic, neven.miculinic, stipan.mikulic\}@fer.hr}\\
}

@@

% \begin{table}[h!]
% \caption{STS score description, adapted from~\citep{agirre2016semeval}}
% \label{tab:sts-score}
% \begin{center}
% 	\resizebox{\columnwidth}{!}{
% \begin{tabularx}{\columnwidth}{llr}
% \toprule
% Score & Explanation\\
% \midrule
% 5 & \textit{Two sentences are completely equivalent}\\
% & The bird is bathing in the sink\\
% & Birdie is washing itself in the water basin.\\
% \midrule
% 4 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
%   & \multicolumn{1}{m{6.5cm}}{In May 2010, the troops attempted to invade
% Kabul.} \\
% & \multicolumn{1}{m{6.5cm}}{The US army invaded Kabul on May 7th last
% year, 2010.} \\
% \midrule
% 3 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
%   & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
% suspect.}\\
% & ``He is not a suspect anymore.'' John said.\\
% \midrule
% 2 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
% &They flew out of the nest in groups. \\
% &They flew into the nest together. \\
% \midrule
% 1 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
% & The woman is playing the violin.\\
% & The young lady enjoys listening to the guitar.\\
% \midrule
% 0 & \textit{The two sentences are completely dissimilar.}\\
%   & \multicolumn{1}{m{6.5cm}}{John went horse back riding at dawn with a whole group of friends.}\\
%   & \multicolumn{1}{m{6.5cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
% \bottomrule
% \end{tabularx}
% }
% \end{center}
% \end{table*}
% 
\section{Related Work}

@@

\section{Extent of the Paper}

@@

\subsection{Data Analysis}
We used dataset from SemEval 2017
Task 1\footnote{http://alt.qcri.org/semeval2017/task1/}.
The dataset has 250 instances. Each instance consists of two sentences and our tash is to predict semantic similarity between them. In order to get some deeper insights on data we did some minor data analysis. We got to conclusion that we are dealing with very short text. Also, we concluded that the smaller difference between lengths of sentences is, the more similar they are. This can be seen clearer on following plots.

\begin{table}
\caption{Stats about lengths of sentences.}
\label{tab:narrow-table}
\begin{center}
\begin{tabular}{cll}
\toprule
& Without stopwords & With stopwords \\
\midrule
Sentence length & 33.004 & 43.838 \\
Tokens in sentence & 5.488 & 8.702 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{sen_len_no_stop.png}
%\caption{Sentence lengths without stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{sen_len_with_stop.png}
%\caption{Sentence lengths with stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{tokens_no_stop.png}
%\caption{Number of tokens without stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
%
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=\columnwidth]{tokens_with_stop.png}
%\caption{Number of tokens with stopwords}
%\label{fig:lstm_2nd_layer}
%\end{center}
%\end{figure}
\subsection{Baseline Models}

@@

\subsubsection{Jaccard Similarity Baseline}

@@

\subsubsection{Cosine Similarity}

@@

\begin{table}
\caption{Averaged scores.}
\label{tab:narrow-table}
\begin{center}
\begin{tabular}{ccc}
\toprule
& R2 & Pearson\\
\midrule
score & 0.5898 & 0.7795 \\
std & 0.011393 & 0.006625 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\subsection{Deep Learning}

@@

\subsubsection{Recurrent Networks}

@@

\begin{table}
\caption{Sample network output.}
\label{tab:sample_output}
\begin{center}
\begin{tabular}{c}
\toprule
Sample output of the LSTM \\
\midrule
\multicolumn{1}{m{6.5cm}}{Little Joe Otter had had no honour pointing to help him so that they
heard of his own fish.  And it was for the beginning of the Little Fir Tree and
scrambled to the sheep. 
Come, let us go and let the right out of this tree, and the LORD had chanced
to the next tabernacce of Mischians, when the king was carried the short of the children of the Green.}
\bottomrule
\end{tabular}
\end{center}
\end{table}

@@

\subsection{All Results}

@@

\begin{table}
	\caption{Model comparison.}
\label{tab:all_results}
\begin{center}
\begin{tabular}{ccc}
\toprule
Model & Pearson coefficient \\
\midrule
Jaccard similarity & 0.73 \\
Cosine baseline & 0.73 \\
LSTM & 0.61 \\
SVM & 0.78 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
